---
title: Plants in Danger
author: ~
date: '2020-12-09'
slug: plants-in-danger
categories: []
tags: []
---

```{r, message = FALSE}
library(tidyverse)
library(tinytex)
library(here)
library(sf)
library(ggplot2)
library(tmap)
library(tmaptools)
library(leaflet)
library(dplyr)
```

```{r, message= FALSE}
dat3<- read_csv(here::here("data", "tidytuesday", "data", "2020", "2020-08-18", "plants.csv"
))
dat4<- read_csv(here::here("data", "tidytuesday", "data", "2020", "2020-08-18", "threats.csv"
))
dat5<- read_csv(here::here("data", "tidytuesday", "data", "2020", "2020-08-18", "actions.csv"
))


```

Plants in Danger
For this project, I am examining plants in danger within the data set plants.csv file from the Tidy Tuesday website. The data set has 500 observations of 24 variables. The variable bionomial_name is a factor of 500 levels that represents the species name (Genus + species). The variable country is a factor of 500 levels that represents the country the plant is origin to. The variable continent is a factor of 500 levels that represents the continent the plant originated. The variable group is a factor of 500 levels that represents the taxonomic group of the plant.The variable year_last_seen is a factor of 500 levels that represents the last year the plant was last seen. The next 12 variables all have 500 levels and describe the different threats of the plants (AA - agriculture & aquaculture, BRU- biological resource use, RDC-commercial development, ISGD -invasive species, EPM - energy production and mining, CC- climate change, HID -human intrusions, P- pollution, TS- transportation corridor, NSM- natural systems modifications, GE- geological events, NA- unknown).

The next data set, threats.csv, has 6,000 observations and 5 variables. The variable bionomial_name is factor of 6000 levels that represent the species name (Genus + species). The variable country is factor of 6000 levels that represent the species original location.The variable continent is factor of 6000 levels that represent the species originating continent. The variable group is factor of 6000 levels that represent the taxonomic group the plant belongs to. The variable year_last_seen is factor of 6000 levels the last calendar year the plant was seen.The variable red_list_category is factor of 6000 levels that represent the IUCN Red List category. The variable threat_type is factor of 6000 levels that represent the type of threat the plants are in, from the last data set. The variable threatened is factor of 6000 levels that represent a binary of 0 (action not taken) or 1 (aciton taken).

The last data set, actions.csv has 3000 observations and 8 variables.
The variable bionomial_name is factor of 3000 levels that represent the species name (Genus + species). The variable country is factor of 3000 levels that represent the species original location.The variable continent is factor of 3000 levels that represent the species originating continent. The variable group is factor of 3000 levels that represent the taxonomic group the plant belongs to. The variable year_last_seen is factor of 3000 levels the last calendar year the plant was seen.The variable red_list_category is factor of 3000 levels that represent the IUCN Red List category. The variable action_type is factor of 3000 levels that represent the type of action the plant has had.The variable action_taken is factor of 3000 levels that represent a binary of 0 (action not taken) or 1 (action taken).

```{r}
glimpse(dat3)
```
```{r}
glimpse(dat4)

```
Question 1:
 Which countries in North America currently have the most plants in danger by extinction?
 
To Determine, filter within the data set, observe at a closer lense to the North American Countries
```{r}
filter(dat4, continent %in% c("North America"))%>%
  filter(year_last_seen == c("1980-1999","2000-2020"))%>%
  filter(threat_type == "Pollution")%>%
  filter(red_list_category %in% c("Extinct", "Extinct in the Wild"))
```
Of the plants that became extinct in 2000-2020, the majority were last seen in America. 16 were from the United States, 2 from Mexico, and only 1 from Panama.Therefore, the United States has a greater pollution problem than the other two countries.  



Question 2: Is there trend in the threat type that came help determine the action type that needs to be enforced?
To determine, filter the data to match what the threat was and whether or not action was taken to deter that threat. 
```{r}

```

Question 3: Was there an event in recent history that wiped out, or put multiple plant species in danger? 
To determine, I will create a graphic graphing the year the plants were last seen as well as the threat the plants are/were in. 
```{r}
glimpse(dat5)

```
