---
title: Plants in Danger
author: ~
date: '2020-12-09'
slug: plants-in-danger
categories: []
tags: []
---

```{r, message = FALSE, echo = TRUE}
library(tidyverse)
library(tinytex)
library(here)
library(sf)
library(ggplot2)
library(tmap)
library(tmaptools)
library(leaflet)
library(dplyr)
```

```{r, message= FALSE,echo = TRUE}
dat3<- read_csv(here::here("data", "tidytuesday", "data", "2020", "2020-08-18", "plants.csv"
))
dat4<- read_csv(here::here("data", "tidytuesday", "data", "2020", "2020-08-18", "threats.csv"
))
dat5<- read_csv(here::here("data", "tidytuesday", "data", "2020", "2020-08-18", "actions.csv"
))


```

## Plants in Danger
For this project, I am examining plants in danger within the data set plants.csv file from the Tidy Tuesday website. The data set has 500 observations of 24 variables. The variable bionomial_name is a factor of 500 levels that represents the species name (Genus + species). The variable country is a factor of 500 levels that represents the country the plant is origin to. The variable continent is a factor of 500 levels that represents the continent the plant originated. The variable group is a factor of 500 levels that represents the taxonomic group of the plant.The variable year_last_seen is a factor of 500 levels that represents the last year the plant was last seen. The next 12 variables all have 500 levels and describe the different threats of the plants (AA - agriculture & aquaculture, BRU- biological resource use, RDC-commercial development, ISGD -invasive species, EPM - energy production and mining, CC- climate change, HID -human intrusions, P- pollution, TS- transportation corridor, NSM- natural systems modifications, GE- geological events, NA- unknown).

The next data set, threats.csv, has 6,000 observations and 5 variables. The variable bionomial_name is factor of 6000 levels that represent the species name (Genus + species). The variable country is factor of 6000 levels that represent the species original location.The variable continent is factor of 6000 levels that represent the species originating continent. The variable group is factor of 6000 levels that represent the taxonomic group the plant belongs to. The variable year_last_seen is factor of 6000 levels the last calendar year the plant was seen.The variable red_list_category is factor of 6000 levels that represent the IUCN Red List category. The variable threat_type is factor of 6000 levels that represent the type of threat the plants are in, from the last data set. The variable threatened is factor of 6000 levels that represent a binary of 0 (action not taken) or 1 (aciton taken).

The last data set, actions.csv has 3000 observations and 8 variables.
The variable bionomial_name is factor of 3000 levels that represent the species name (Genus + species). The variable country is factor of 3000 levels that represent the species original location.The variable continent is factor of 3000 levels that represent the species originating continent. The variable group is factor of 3000 levels that represent the taxonomic group the plant belongs to. The variable year_last_seen is factor of 3000 levels the last calendar year the plant was seen.The variable red_list_category is factor of 3000 levels that represent the IUCN Red List category. The variable action_type is factor of 3000 levels that represent the type of action the plant has had.The variable action_taken is factor of 3000 levels that represent a binary of 0 (action not taken) or 1 (action taken).

```{r}
glimpse(dat3)
```

```{r}
glimpse(dat4)
```

# Question 1
 Which countries in North America have the most disappearances of Plants in the years 2000-2020 due to pollution?
 
To Determine, filter within the data set, observe at a closer lense to the North American Countries using filters to get only the variables I wish to see.
```{r}
q1<-filter(dat4, continent %in% c( "North America"))%>%
  group_by(country)%>%
  filter(year_last_seen == c("2000-2020"))%>%
  filter(threat_type == "Pollution")%>%
  ggplot(aes(x = country, count = threat_type, fill = threat_type))+
  geom_bar()+
  ggtitle("Pollution reason Plants Disappear in the years 2000-2020 in North America")+
  ylab("Number Disappeared")+
  xlab("Country in North America")
q1



```
Of the plants that became extinct in 2000-2020 due to pollution, the majority were last seen in America. 16 were from the United States, 2 from Mexico, 1 from Panama. This insinuates that America has a much larger pollution problem at hand and it is detrimental to their plant species. 


If we take it further to examine which countries in North America are most fatal to their plants we can analyze the following:
```{r}
q1b<-filter(dat4, continent %in% c( "North America"))%>%
  group_by(country)%>%
  filter(threat_type == "Pollution")%>%
  ggplot(aes(x = country, count = threat_type, fill = threat_type))+
  geom_bar()+
  ggtitle("Pollution reason Plants Disappear in North America")+
  ylab("Number Disappeared")+
  xlab("Country in North America")+
  theme(axis.text.x = element_text(angle =- 45, hjust = 0))
q1b

  
```
This graph portrays the amount of plants each country has lost, not filtered by the current years. This shows that America is consistently the most harmful to their plants due to pollution than any other North American country or Island. 

If we examine it even further and look at which countries are more harmful in general (not just looking at pollution like we have in the first two graphs). We should analyze the following:
```{r}
q1c<-filter(dat4, continent %in% c( "North America"))%>%
  group_by(country)%>%
  ggplot(aes(x = country, count = threat_type, fill = threat_type))+
  geom_bar()+
  ggtitle("Plants Disappear in North America")+
  ylab("Number Disappeared")+
  xlab("Country in North America")+
  theme(axis.text.x = element_text(angle =- 45, hjust = 0))
q1c

  
```
The distributions look very similar to the second graph in the first question, however, looking at the units on the y axis, it is apparent that this graph displays the total of disappeared plants in North America. And like how the other graphs displayed, the United States again is the most harmful to plants, no matter what the threat type is. 


# Question 2 
How much action is being enforced for the plants in North American Countries?
To determine, filter the data to match what the threat was and whether or not action was taken to deter that threat. 
```{r}
action<- filter(dat5, action_taken == "1")%>%
  filter(year_last_seen == "2000-2020")

action_total <- filter(dat5 ,action_taken == c("0","1")) %>%
  filter(year_last_seen == c("2000-2020"))


proportion <- count(action)/count(action_total)

proportion

action_in_NA<- filter(dat5, action_taken == "1")%>%
  filter(continent %in% c("North America"))%>%
  filter(year_last_seen == "2000-2020")

action_total_NA <- filter(dat5 ,action_taken == c("0","1")) %>%
  filter(continent %in% c("North America"))%>%
  filter(year_last_seen == c("2000-2020"))

proportion_NA <- count(action_in_NA)/count(action_total_NA)
proportion_NA

```
America actually has a slightly greater ratio of action taken to action not taken than the world in total. This makes sense as the United States seems to have one of the greatest extinction rates in the world.



# Question 3
Is there a year interval where there is a multitude of plants lost in the United States? If so, what is the likely cause for such disappearance?
```{r}
dat4%>%
  filter(country == "United States")%>%
  filter(binomial_name == c("Acaena exigua","Achyranthes atollensis","Adenophorus periens","Brighamia insignis","Clermontia multiflora",  "Stenogyne kanehoana"))%>%
  ggplot(aes(y= binomial_name, x = year_last_seen))+
  geom_point(mapping = aes(y= binomial_name, x = year_last_seen))+
  theme(axis.text.x = element_text(angle =- 45, hjust = 0))+
  xlab("Year Last Seen") +
  ylab("Species Name") +
  ggtitle("Year Last Seen for Plant Species in United States")


```
To evaluate this question, I picked random plants that were last seen in the United States and compared their disappearance. My reasoning for picking random points is that an event that would "wipe out" different species would not value one type of species above the other, so I figured picking random plants would create an understanding of such event. Out of the 6 I picked, 1/2 of them were last seen in the years 2000-2020. This is likely due to the harm that a more urbanized society has on plant life, such as pollution.
```{r}
dat4%>%
  filter(country == "United States")%>%
  filter(binomial_name == c("Acaena exigua","Achyranthes atollensis","Adenophorus periens","Brighamia insignis","Clermontia multiflora",  "Stenogyne kanehoana"))%>%
  filter(year_last_seen == "2000-2020")


```
In response to my last claim that it is likely pollution that is so harmful to these plants, not one of them from the group I selected above disappeared from pollution. In fact, all of the 6 plants had a different reason for disappearance. The closest to my hypothesis was that plant Adenophorus periens was threatened by commercial development, the urbanization of America.


# Conclusion
This data set provides very valuable information about the future of plants. This becomes a human problem because we are so dependent on plants to produce the oxygen in our atmosphere. It is apparent in almost all the data sets that America is the leading "killer" of plants, however, they have had enacted action to deter this fact. The loss of plants is inevitable, especially with increased urbanization and the negative side effects that it brings (climate change, pollution, commercial development, etc). But as urbanization continues, so must the action taken by local or federal government to ensure the safety of these valuable plants. 